Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory (HWO), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
%
Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
%
The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
%
Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
%
We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
%
We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
%
The glint signature at planetary phase angles of $\sim$50--70$^\circ$ 
%\sim$110--130$^\circ$ 
would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA{}; 3$\lambda/D$ for a 6-m telescope at 600nm).
%
Increasing the \IWA{} to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
%
The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.    \end{abstract}

@@

In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds (HWO) target stars.
%
In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the range of phase angles given a few different limits, in \S 2.3, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4, we outline how the planet/star contrast changes as a function of the planet's orbital phase.  
%
We present and discuss our results in \S 3 and conclude in \S 4. It is important to note that the phases for each feature throughout this paper are approximate, that features overlap, and that the maximum peak can vary with composition and wavelength.
%the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.

@@

\begin{table}
    \centering
    \caption{
        The ranges of phase angles for the different scattering phenomena considered here.
    }
    \label{tab:phase_ranges}
    \begin{tabular}{ c c c c } 
    \toprule
     & $\varphi_{min}$ & $\varphi_{peak}$ & $\varphi_{max}$ \\
    \midrule
    \midrule
    Rainbows & 22 & 42 & 63 \\
    Rayleigh & 50 & 70 & 110 \\
    Ocean Glint & 130 & 150 & 170 \\
    Glory & 0 & 5 & 10 \\
    \bottomrule
    \end{tabular}
\end{table}

>>>>

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
%
Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
%
The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
%
Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
%
We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
%
We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
%
The glint signature at planetary phase angles of $\sim$50--70$^\circ$ 
%\sim$110--130$^\circ$ 
would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA{}; 3$\lambda/D$ for a 6-m telescope at 600nm).
%
Increasing the \IWA{} to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
%
The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.    \end{abstract}

@@

In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
%
In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the range of phase angles given a few different limits, in \S 2.3, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4, we outline how the planet/star contrast changes as a function of the planet's orbital phase.  
%
We present and discuss our results in \S 3 and conclude in \S 4. It is important to note that the phases for each feature throughout this paper are approximate, that features overlap, and that the maximum peak can vary with composition and wavelength.
%the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.

@@

\begin{table}
    \centering
    \caption{
        The ranges of phase angles for the different scattering phenomena considered here.
    }
    \label{tab:phase_ranges}
    \begin{tabular}{ c c c c } 
    \toprule
     & $\varphi_\text{min}$ & $\varphi_\text{peak}$ & $\varphi_\text{max}$ \\
    \midrule
    \midrule
    Rainbows & 22 & 42 & 63 \\
    Rayleigh & 50 & 70 & 110 \\
    Ocean Glint & 130 & 150 & 170 \\
    Glory & 0 & 5 & 10 \\
    \bottomrule
    \end{tabular}
\end{table}
